---
title: "Census ARM"
output: html_notebook
---


```{r setup}
census_data <- read.csv("ACS_Full_Features.csv")
head(census_data)
```
It seems the default column names aren't in the right format, let's fix that.

```{r}

colnames(census_data) <- gsub("\\.", " ", colnames(census_data))
print(colnames(census_data))
```

Now that that's done, time to drop the unnecessary columns and check if any
others need to be binned.

Age and Marriage aren't binned, 

```{r}
describe(census_data$`Median Age by Sex`)

```
The minimum is 21, the max is 68, so under 28 will be 'Young', 28-35 will be
'Adult', 35-45 will be 'Mature Adult', 45-55 will be 'Older Adult', and 55+
will be 'Senior'.

```{r}
breaks = c(0, 28, 35, 45, 55, Inf)
labels = c('Young', 'Adult', 'Mature Adult', 'Older Adult', 'Senior')
```

```{r}
census_data <- census_data %>% 
  mutate(Age_Category = cut(`Median Age by Sex`, breaks = breaks,
                            labels = labels, right = F))
```

```{r}
table(census_data$Age_Category)

```

Not much of a 'spread' in this data, but that's alright.
```{r}
describe(census_data$`Now Married`)
```

Marriage seems harder to bin, given that it is based on population, with a vast
spread. 0, 2000, 6000, 8000, 20000, and 100000 will be the bin numbers,
with No, Few, Low, Median, Many, and High being the categories.

```{r}
breaks2 <- c(0, 2000, 6000, 8000, 20000, 100000, Inf)
labels2 <- c('No Marriages', 'Few Marriages', 'Low Marriages', 'Median Marriages',
             'Many Marriages', 'High Marriages')
```

```{r}
census_data <- census_data %>% 
  mutate(Marriage_Category = cut(`Now Married`, breaks = breaks2,
                            labels = labels2, right = F))
```

```{r}
table(census_data$Marriage_Category)
```

Good spread! Now, it's time to discard the numeric categories, and the NAME,
state, and county categories. Income Category will be left off, because it's
the label for the data.

```{r}
census_arm <- census_data %>% select(Age_Category, Marriage_Category,
                     `POC Population Category`, `Poverty Rate Classification`,
                     `Disabled Population Category`)
arm_label <- census_data %>% select(`Income Category`)
```

```{r}
head(census_arm)
```
```{r}
write.csv(census_arm, "C:\\Users\\gdehl\\OneDrive\\Documents\\College -\\Grad School\\Machine Learning\\ARM_Census.csv")
```

```{r}
head(arm_label)
```

Now, to turn this into transaction data.

```{r}
census_arm <- census_arm %>% mutate_if(is.character, as.factor)

sapply(census_arm, class)
```


Next comes converting the factor data into transactions.
```{r}
census_basket <- as(census_arm, 'transactions')
inspect(head(census_basket, n = 5))
```

Finally, the apriori algorithm can be applied to the data.
```{r}
census_rules <- apriori(census_basket, 
                        parameter = list(support = 0.2, 
                                         confidence = 0.6))

# Inspect the rules
inspect(census_rules)
```
The empty first two rules indicate a precedent for 'Mature Adult' and
'Low Poverty' to appear in all 'transactions' in the dataset, or they appear so
frequently that they count all on their own. Filtering them out to get better
results is a good idea:

```{r}
filtered_rules <- apriori(census_basket, 
                        parameter = list(support = 0.2, 
                                         confidence = 0.6, minlen = 2))
inspect(filtered_rules)
```

One interesting thing noted is the low lift rates on some of the top rules,
a number of them have less than 1 for their lift.

Visualization comes last.
```{r}
top_rules <- head(sort(filtered_rules, by = "lift"), 15)
inspect(top_rules)
p1 <- plot(top_rules, method = "graph", engine = "htmlwidget")
p1
```

```{r}
top_rules2 <- head(sort(filtered_rules, by = 'confidence'), 15)
inspect(top_rules2)
p2 <- plot(top_rules2, method = 'graph', engine = 'htmlwidget')
p2
```

```{r}
top_rules3 <- head(sort(filtered_rules, by = 'support'), 15)
inspect(top_rules3)
p3 <- plot(top_rules3, method = 'graph', engine = 'htmlwidget')
p3
```

Finally, time to save the graphs
```{r}
# Use savewidget to save html files for display on website
# Make sure to library (htmlwidgets)
saveWidget(p1, file = "lift_plot.html")
saveWidget(p2, file = "confidence_plot.html")
saveWidget(p3, file = "support_plot.html")
```